Kaggle-in-class Data Challenges can Boost Student Learning

### 2018-03-24 paper-kaggle.Rmd
intro filled in

Kaggle is well-known for the richly funded data competitinns, where participants compete to score the lowest error in their model fitting. Recently, they have made it possible to run in-class competitions, private and limited to members of a university course. This work explores how student participation in these challenges improves performance, engagement, interest

@@

Explain experimental design: two compeitions, two locations, two courses, one with students randomised into two groups, the other with control group of students and regression challenge, and exam questions. (Not sure how to incorporate ETC3250 students at the moment.)

@@

-	To examine different type of students (undergraduate vs post graduate) and different learning environments (different university) the study run across three different subjects and two universities, Monash University (ETC3250 and ETC5242) and University of Melbourne (MAST90083 subject).

-	Time period: semester two 2017

-	MAST90083:

63 enrolled students; postgraduate level; background: math & stats, IT and actuarial science;

@@

Note that the data competition was compulsory to all the students. The volunteering part was the questionnaire at the end of the semester asking for feedback on aspects of the class such as their competition experience, their favourite topic of the semester, and their enjoyment of the material in the subject.

-	ETC5242

-	ETC3250

## Competitions

## Used Platform

Kaggle in class (https://inclass.kaggle.com)...

@@

```{r}
# read Unimelb data
UniMelbData <- read.csv(file="data/UniMeldUnIndef.csv", header=TRUE, sep=",")
```

@@

@@

@@

@@

```{r permtest}
pdif <- NULL
for (i in 1:1000) {
  x <- scores %>% select(class, perf)
  x <- x %>% mutate(class = sample(class))
  pdif <- c(pdif,
            mean(x$perf[x$class=="Yes"]-
                   mean(x$perf[x$class=="No"])))
}
dif <- mean(scores$perf[scores$class=="Yes"]-
       mean(scores$perf[scores$class=="No"]))
```
```{r}
tt <- t.test(scores$perf[scores$class=="Yes"], scores$perf[scores$class=="No"], alternative="greater")
```

On average the students who participated in the kaggle challenge scored `r dif` higher than those that didn't, `r mean(scores$perf[scores$class=="Yes"])` in comaprison to `r mean(scores$perf[scores$class=="No"])`. Using a permutation t-test, this corresponds to a significant difference in means, with $p$-value of `r length(pdif[pdif >= dif])/1000`. A conventional $t$-test, yields similar results with a $p$-value of `r tt$p.value`.

@@

This project (title: Effect of Data Competition on Learning Experience) has been approved by the Faculty of Science Human Ethics Advisory Group University of Melbourne (Ethics ID: 1749858.1 on September the 4th, 2017) and by XXX Ethics Advisory Group Monash University (Ethics ID: XXX on XXX, 2017).

>>>>

Kaggle [@kaggle] is well-known for the data competitions, some richly funded. It provides a platform for predictive modelling and analytics competitions where participants compete to produce the best predictive model for a given data set. In 2015, Kaggle InClass was introduced, as a self-service platform to conduct competitions. These competitions can be private, limited to members of a university course, and are easy to setup. This paper examines the educational benefits of conducting predictive modeling competitions in class on performance, engagement and interest.

@@

The experiment was conducted during Fall semester 2017. Data was collected during three classes, one at the University of Melbourne (MAST90083), and two at Monash University (ETC2420/5242 and ETC3250).

## Competition data

Two data sets were compiled for the kaggle challenges: Melbourne property auction prices and spam classification. The Melbourne auction price data was compiled by extracting information from real estate auction reports (pdf) collected between Feb 2, 2013 and Dec 17, 2016. Students were expected to predict price based on the property characteristics. The spam classificaation data was compiled by graduate students at Iowa State University as part of a statistical computing class by Dr Heike Hofmann, in XXX. Data was compiled by monitoring and extracting information from emails over a period of a week, and manually classifying them as spam or ham. Students were expected to classify the email as spam or not.

Both data sets provide substantial challenge for prediction.

@@

MAST90083 is titled Computational Statistics and Data Mining, is designed for postgraduate level, for students with math, statistics, information technology or actuarial backgrounds. It covers modelling both continuous (regression) and categorical (classification) response variables. The 63 students were randomized into one of two kaggle competitions, one focused on regression (R) and the other classification (C). Students individually built prediction models and made submissions for 16 days, and then were allowed to form groups to compete for another 7 days.

ETC2420/5242, titled Statistical Thinking, covers regression, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate students were required to participate in the kaggle competition focused on regression (R). The 145 undergraduate students are considered control for examining performance. The competition ran for one month. Students formed their own teams of 2-4 members to compete. Several undergraduates also chose to compete individually.

ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on a classification problem. Because this group had no comparison group, it was difficult to assess performance. This data was primarily used to examine engagement and interest based on a follow-up questionnaire.

<!--
63 enrolled students; postgraduate level; background: math & stats, IT and actuarial science;

@@

Note that the data competition was compulsory to all the students. The volunteering part was the questionnaire at the end of the semester asking for feedback on aspects of the class such as their competition experience, their favourite topic of the semester, and their enjoyment of the material in the subject.
-->

## Platform

MAST90083 used \url{https://inclass.kaggle.com/c}. ETC2420/5242 used \url{https://inclass.kaggle.com/c/vitticeps}. ETC3250 used \url{}.

@@

```{r}
# read Unimelb data
# UniMelbData <- read_csv(file="data/UniMeldUnIndef.csv")
# UniMelbData <- UniMelbData %>% select(-LastName, -FirstName, -StudentID)
# write_csv(UniMelbData, path="data/deidentified_UMelb.csv")
UniMelbData <- read_csv(file="data/deidentified_UMelb.csv") #, header=TRUE, sep=",")
```

@@

```{r eval=FALSE}
perm_test <- function(x, cl, n){ # x- table: 1col proj_alloc, 2col perf (relevant score)
  # Original data
  mean_reg   <- median(x$perf[x$proj_alloc=="Melbourne Price"])
  mean_class <- median(x$perf[x$proj_alloc=="Spam classification"])
  dif        <- median(x$perf[x$proj_alloc=="Melbourne Price"]-
                  median(x$perf[x$proj_alloc=="Spam classification"]))

  # permutation t-test
  pdif <- NULL
  for (i in 1:n) {
  x <- x %>% mutate(proj_alloc = sample(proj_alloc))
  pdif <- c(pdif,
            median(x$perf[x$proj_alloc=="Melbourne Price"]-
                   median(x$perf[x$proj_alloc=="Spam classification"])))
  }
  pv_perm <- length(pdif[pdif >= dif])/n

  # conventional t-test
  pv_ttest <- t.test(x$perf[x$proj_alloc=="Melbourne Price"], 
                     x$perf[x$proj_alloc=="Spam classification"], 
                     alternative="greater")$p.value

  return(list("mean_reg"= mean_reg, "mean_class"= mean_class, "dif" = dif, "pv_perm" = pv_perm, "pv_ttest" = pv_ttest))
}  
```

```{r echo = FALSE, eval=FALSE}
num_permutations = 1000
set.seed(123456)# fix seed

#PTQ
x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_class)
perm_test_ptq_class <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_reg)
perm_test_ptq_reg <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_class_reg)
perm_test_ptq_class_reg <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_noncompq)
perm_test_ptq_noncompq <- perm_test(x = x, n = num_permutations)

### PTE
x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_class)
perm_test_pte_class <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_reg)
perm_test_pte_reg <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_class_reg)
perm_test_pte_class_reg <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_noncompq)
perm_test_pte_noncompq <- perm_test(x = x, n = num_permutations)

## preper a table

tt <- cbind(
t( cbind(perm_test_ptq_class , perm_test_ptq_reg , perm_test_ptq_class_reg , perm_test_ptq_noncompq)),

c("","","",""),

t( cbind(perm_test_pte_class , perm_test_pte_reg , perm_test_pte_class_reg , perm_test_pte_noncompq))
)
tt <- rbind( c("PTQ" , "PTQ" , "PTQ", "PTQ" , "PTQ" ,"" , "PTE", "PTE", "PTE", "PTE", "PTE") , tt)
tt <- rbind(c("Mean MelbP", "Mean Spam" , "Diff", "P.v. Perm", "P.v. t-test" , "",
                  "Mean MelbP", "Mean Spam" , "Diff", "P.v. Perm", "P.v. t-test"), tt)

tt <- cbind( c("", "" , "Classification", "Regression", "Class & Reg", "Other"), tt)
data.frame(tt)

```{r}
pdif <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_class)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif <- c(pdif,
            median(x$pte_class[x$class=="Spam classification"]-
                   median(x$pte_class[x$class=="Melbourne Price"])))
}
dif <- median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="Spam classification"]-
       median(UniMelbData$pte_class[UniMelbData$Proj_alloc=="Melbourne Price"]))
length(pdif[pdif >= dif])/1000
```

```{r}
pdif <- NULL
for (i in 1:1000) {
  x <- UniMelbData %>% select(Proj_alloc, pte_reg)
  x <- x %>% mutate(class = sample(Proj_alloc))
  pdif <- c(pdif,
            median(x$pte_reg[x$class=="Melbourne Price"]-
                   median(x$pte_reg[x$class=="Spam classification"])))
}
dif <- median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="Melbourne Price"]-
       median(UniMelbData$pte_reg[UniMelbData$Proj_alloc=="Spam classification"]))
length(pdif[pdif >= dif])/1000
```

Next we examined the statistical significance of the mean differences between the performances of the two group of students.

@@

```{r permutation_class_questions, echo = FALSE, eval=FALSE}

set.seed(123456)# fix seed

#PTQ
# x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_class_Q1) # 0,1 question
# perm_test_ptq_class_Q1 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_class_Q10)
perm_test_ptq_class_Q10 <- perm_test(x = x, n = num_permutations)

#PTE
# x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_class_Q1)
# perm_test_pte_class_Q1 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_class_Q10)
perm_test_pte_class_Q10 <- perm_test(x = x, n = num_permutations)

## preper a table

tt <- cbind(
t(  perm_test_ptq_class_Q10),

c(""),

t( perm_test_pte_class_Q10 )
)

tt <- rbind( c("PTQ" , "PTQ" , "PTQ", "PTQ" , "PTQ" ,"" , "PTE", "PTE", "PTE", "PTE", "PTE") , tt)
tt <- rbind(c("Mean MelbP", "Mean Spam" , "Diff", "P.v. Perm", "P.v. t-test" , "",
                  "Mean MelbP", "Mean Spam" , "Diff", "P.v. Perm", "P.v. t-test"), tt)

tt <- cbind( c("", "", "Q10"), tt)
data.frame(tt)

Next we examined the statistical significance of the mean differences between the performances of the two group of students for Q10.

@@

```{r permutation_Reg_questions, echo = FALSE, eval=FALSE}

set.seed(123456)# fix seed

#PTQ
x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_reg_Q6)
perm_test_ptq_reg_Q6 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_reg_Q8)
perm_test_ptq_reg_Q8 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_reg_Q15)
perm_test_ptq_reg_Q15 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = ptq_reg_Q16)
perm_test_ptq_reg_Q16 <- perm_test(x = x, n = num_permutations)

### PTE
x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_reg_Q6)
perm_test_pte_reg_Q6 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_reg_Q8)
perm_test_pte_reg_Q8 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_reg_Q15)
perm_test_pte_reg_Q15 <- perm_test(x = x, n = num_permutations)

x <-  UniMelbData %>% select( proj_alloc = Proj_alloc, perf = pte_reg_Q16)
perm_test_pte_reg_Q16 <- perm_test(x = x, n = num_permutations)

## preper a table

tt <- cbind(
t( cbind(perm_test_ptq_reg_Q6 , perm_test_ptq_reg_Q8, perm_test_ptq_reg_Q15 , perm_test_ptq_reg_Q16)),

c("","","",""),

t( cbind(perm_test_pte_reg_Q6 , perm_test_pte_reg_Q8 , perm_test_pte_reg_Q15 , perm_test_pte_reg_Q16))
)
tt <- rbind( c("PTQ" , "PTQ" , "PTQ", "PTQ" , "PTQ" ,"" , "PTE", "PTE", "PTE", "PTE", "PTE") , tt)
tt <- rbind(c("Mean MelbP", "Mean Spam" , "Diff", "P.v. Perm", "P.v. t-test" , "",
                  "Mean MelbP", "Mean Spam" , "Diff", "P.v. Perm", "P.v. t-test"), tt)

tt <- cbind( c("", "" , "Q6", "Q8", "Q15", "Q16"), tt)
data.frame(tt)

Next we examined the statistical significance of the mean differences between the performances of the two group of students for Q6, Q8, Q15 and Q16.

@@

```{r permtest}
pdif <- NULL
for (i in 1:1000) {
  x <- scores %>% select(class, perf)
  x <- x %>% mutate(class = sample(class))
  pdif <- c(pdif,
            median(x$perf[x$class=="Yes"]-
                   median(x$perf[x$class=="No"])))
}
dif <- median(scores$perf[scores$class=="Yes"]-
       median(scores$perf[scores$class=="No"]))
```
```{r}
tt <- t.test(scores$perf[scores$class=="Yes"], scores$perf[scores$class=="No"], alternative="greater")
```

On average the students who participated in the kaggle challenge scored `r dif` higher than those that didn't, a median of `r mean(scores$perf[scores$class=="Yes"])` in comparison to `r mean(scores$perf[scores$class=="No"])`. Using a permutation test, this corresponds to a significant difference in medians, with $p$-value of `r length(pdif[pdif >= dif])/1000`. A conventional two-sample $t$-test, of mean difference, yields similar results with a $p$-value of `r tt$p.value`.

@@

This project (title: Effect of Data Competition on Learning Experience) has been approved by the Faculty of Science Human Ethics Advisory Group University of Melbourne (ID: 1749858.1 on September the 4th, 2017) and by Monash University Human Research Ethics Committee (ID: 9985 on 24/08/2017).
